The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games

### 2025-05-06 Quarto/paper.qmd
Fix citations in Methods and Introduction

- We wanted to compare how humans and large language models (LLMs) handle these situations. To do this, we used a recent human dataset and constructed one from an LLM-based simulator: 
	- _Werewolf Among Us_ (Lai et al., 2022), a collection of real human gameplay annotated with persuasion strategies, 
	- _Werewolf Arena_ (Bailis et al., 2024), a simulated environment where LLM agents play the game autonomously. 
	- While both studies show Werewolf generates complex strategic language, neither compares human and LLM behavior directly.

@@

We used the _Werewolf Among Us_ dataset (Lai et al., 2022), which contains annotated dialogues from over 150 real games of One Night Werewolf and Avalon. These games differ from classic Werewolf because they have only one round of discussion and voting, do not eliminate players during the game, and include many specialized roles beyond Villager and Werewolf. The dataset provides detailed annotations of persuasion strategies for each utterance, such as accusations, defenses, and identity claims. We specifically used the textual transcriptions and their strategy annotations for our comparisons.

@@

The _Werewolf Arena_ dataset (Bailis et al., 2024) features simulated classic Werewolf games played by LLM agents without human intervention. Unlike the one-round human games, these simulations involve multiple rounds alternating between night (secret actions) and day (open discussion). Each LLM agent is assigned a role (Villager, Werewolf, Seer, Doctor) and receives context-specific prompts that guide their strategic interactions.

>>>>

- We wanted to compare how humans and large language models (LLMs) handle these situations. To do this, we used a recent human dataset and constructed one from an LLM-based simulator: 
	- _Werewolf Among Us_ [@laiWerewolfUsMultimodal2022], a collection of real human gameplay annotated with persuasion strategies, 
	- _Werewolf Arena_ [@bailisWerewolfArenaCase2024], a simulated environment where LLM agents play the game autonomously. 
	- While both studies show Werewolf generates complex strategic language, neither compares human and LLM behavior directly.

@@

We used the _Werewolf Among Us_ dataset [@laiWerewolfUsMultimodal2022], which contains annotated dialogues from over 150 real games of One Night Werewolf and Avalon. These games differ from classic Werewolf because they have only one round of discussion and voting, do not eliminate players during the game, and include many specialized roles beyond Villager and Werewolf. The dataset provides detailed annotations of persuasion strategies for each utterance, such as accusations, defenses, and identity claims. We specifically used the textual transcriptions and their strategy annotations for our comparisons.

@@

The _Werewolf Arena_ dataset [@bailisWerewolfArenaCase2024] features simulated classic Werewolf games played by LLM agents without human intervention. Unlike the one-round human games, these simulations involve multiple rounds alternating between night (secret actions) and day (open discussion). Each LLM agent is assigned a role (Villager, Werewolf, Seer, Doctor) and receives context-specific prompts that guide their strategic interactions.
